Learning Auditory Similarities Between Instrumental Playing Techniques

### 2019-10-10 paper/paper.tex
conclude the related work section

%\begin{figure}
%\caption{\csentence{STRF + scattering (upsampling STRF)}
%A short description of the figure content should go here.}
%\end{figure}

The reasons above make STRFs an appealing feature extractor for a perceptual description of timbral similarity across instrumental playing techniques.
Nonetheless, current implementations of STRF suffer from a lack of scalability, which explains why they have found few applications in MIR thus far.
Indeed, the full cortical model is usually computed via two-dimensional Fourier transforms over adjacent time--frequency regions, followed by averaging around specific rates and scales.
On the contrary, joint time--frequency scattering offers a faster extraction of spectrotemporal modulations while preserving properties of differentiability \cite{andreux2019jmlr} and invertibility \cite{lostanlen2019dafx}.
Such acceleration is made possible by discretizing the wavelet transforms involved in time--frequency scattering according to a multirate scheme, both along the time and the log-frequency variables \cite{anden2019tsp}.
This multirate scheme also has the benefit of drawing an explicit connection between scattering networks and deep convolutional networks, as both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class.
Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be known whether this findings can scale up from a single instrument to an entire orchestra.
In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation in terms of its ability to reproduce subjective dissimilarity judgments, remotely from any classification task.

In summary, our paper strives to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
From the standpoint of MIR, 
Conversely, from the standpoint of music cognition, this paper 
% enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).

%Our model borrows from both aforementioned lines of research in music information retrieval and music psychology.
%Indeed, the joint time--frequency scattering transform can be regarded as a deep convolutional network with fixed synaptic weights; or, alternatively, as an idealized computational model for STRFs \cite{anden2019tsp}.
%We refer to \cite{bellet2015book} for a review of the state of the art in this domain.

>>>>

The reasons above make STRFs an appealing feature extractor for a perceptual description of timbral similarity across instrumental playing techniques.
Nonetheless, current implementations of STRF suffer from a lack of scalability, which explains why they have found few applications in MIR thus far.
Indeed, the full cortical model is usually computed via two-dimensional Fourier transforms over adjacent time--frequency regions, followed by averaging around specific rates and scales.
On the contrary, joint time--frequency scattering offers a faster extraction of spectrotemporal modulations while preserving properties of differentiability \cite{andreux2019jmlr} and invertibility \cite{lostanlen2019dafx}.
Such acceleration is made possible by discretizing the wavelet transforms involved in time--frequency scattering according to a multirate scheme, both along the time and the log-frequency variables \cite{anden2019tsp}.
This multirate scheme also has the benefit of drawing an explicit connection between scattering networks and deep convolutional networks, as both involves numerous convolutions with small kernels, pointwise rectifying nonlinearities, and dyadic pooling operations \cite{mallat2016philtrans}.
Moreover, a quantitative benchmark over Medley-solos-DB has demonstrated that joint time--frequency scattering, unlike purely temporal scattering, outperforms deep convolutional networks in supervised musical instrument classification, even in a relatively large data regime with 500 to 5k samples per class.
Recently, Wang et al.~\cite{wang2019ismir} trained a classifier of playing techniques on separable (\ie{} non-joint) time--frequency scattering features, and achieved state-of-the-art results in continuous performance recordings of the Chinese bamboo flute.
Despite supplying experimental evidence that time--frequency scattering is better suited to playing technique recognition than purely temporal scattering, it remains to be known whether this findings can scale up from a single instrument to an entire orchestra.
In addition, previous publications in time--frequency scattering, whether separable or joint, lacks a human-centric evaluation, remotely from any classification task.

In summary, this paper strives to fill the gap in scholarship between MIR and music cognition approaches to timbre, in the context of computer-assisted spectralist orchestration with extended playing techniques.
From the standpoint of MIR, our model offers a tractable and generic multidimensional representation for timbre similarity, alongside formal guarantees of robustness to elastic deformations in the time--frequency domain, whether random or adversarial.
Conversely, from the standpoint of music cognition, our model offers a scalable and biologically plausible surrogate for stimulus-based collection of acoustic dissimilarity judgments, with the potential of being tailored to subjective preferences.

% enjoys a thirty-fold reduction in dimensionality, while covering a  time span that is four times larger (1000 ms) and an acoustic bandwidth that is also four times larger (0--16 kHz).
%We refer to \cite{bellet2015book} for a review of the state of the art in metric learning.
